---
title: Using RStudio's R-Notebooks for Creating Interactive Content for Statistics
  Courses
author: "Dr. Bekah Selby (ESU - Math and Econ)"
output:
  html_document: default
  html_notebook: default
---




```{r,render=F,include=FALSE,echo=FALSE}
setwd("C:/Users/rselby/Google Drive/GitHub/Econometrics/teaching-expo-2019")
knitr::opts_chunk$set(warning=F,message=F)
```

# Introduction 

## About Me

- Dr. Bekah Selby
- Assistant Professor of Economics at Emporia State University
- Data Nerd by trade (applied econometrician)
- Teach classes ranging from basic statistics to graduate level applied econometrics

## My Problems

- Data can be daunting
- Students feel disenfranchised in mathematics courses
- AKA they dread them 
- How to make statistics "come to life"
- How to emphasize transparency 
- How to reduce my own workload 

## One Solution

- Use technology that combines interactivity with transparency
- Rstudio has a feature called "R-Notebooks"
  - Scripting + Word Processing
  - Includes code with visuals and other output
  - Emphasizes process and description 
  - Great place for students to do homework if R is incorporated as part of the curriculum 
- Also good for the teacher: 
  - Create complementary slides and handouts very easily
  - Homework is output to a .html file which renders on LMS
  - Teachers can download .Rmd source file from submissions

## Today

- Introduce some features of notebooks that I frequently use 
- Show you the basic structure of R-Notebooks so you can get started right away
- Discuss creation of interactive content and how it can be incorporated in LMS

# Introduction to R-Studio's R-Notebook

## The YAML Header 

- The first thing you see in a notebook is a YAML header. 
  - title: "An Exceptional Title"
  - author: "Dr. Such and Such"
  - output: html_notebook
  - date: "October 4, 2019"
- This renders in the output as a title. 
- You can choose lots of formats for the output, here we choose `html_notebook` to use the R-notebook capabilities 

## Purpose of a Notebook

- The R-notebook is a way to include all components of an analysis:
  - Code
  - Output
  - Discussion
- It also has the capability of using interactives because it is rendered in html 
- In classroom assignments, this creates an emphasis on transparency of research and analysis (nothing is done "behind the scenes")

## Markdown: 

Uses markdown syntax to create formatted headers (see above), paragraphs, bulleted lists (this is one!), *font* **emphasis**, [hyperlinks](http://rstudio.com), block quotes, images, and more

Renders math-equations using LaTeX. Example: 
```{r,echo=FALSE}
print("The equation $$Y_t = X_t + \varepsilon_t $$ renders to")
```

$$Y_t = X_t + \varepsilon_t $$

- This is commonly used to write up discussion about the analysis!

## R-Chunks

R-chunks are pieces of code that are included in the place where the output is wanted.

- Included Chunk
```{r}
x<-c("Hello","World")
x
```

- Excluded Chunk 
```{r,echo=FALSE}
x<-c("Hello","World")
x
```

- Excluded Output
```{r,include=FALSE}
x<-c("Hello","World")
x
```


## Inline R Code

- You can also write up code using inline syntax
- If we want to calculate the average of `cars$speed`, we might write 
```{r}
mean(cars$speed)
```

or write "The average is `r mean(cars$speed)`" 

# Creating an Interactive Notebook

## Creating Basic Plots and Tables in R 
First things first, using R to create visuals.

We are going to use data already preinstalled in R called `cars`

1. Create a table containing the first 10 observations from the data set `cars`: 

##

```{r}
head(cars,10)
```

## 

2. Create a scatter plot of the speed and distance in the data set `cars` 

##

```{r}
plot(cars)
```

## 

3. Create a summary statistics table including the mean, standard deviation, and min and max for the variables in the data set `cars`.

##

```{r}
#install.packages("psych")
library(psych)
describe(cars,skew = F,trim = 0,IQR = T)
```

##

4. Create a histogram for the speed of cars in the data set `cars`. 

##

```{r}
hist(cars$speed)
```

## Tutorials and Quizzes 

Suppose we want to create an interactive tutorial where students can test their knowledge of R commands. 

```{r,warning=FALSE,message=FALSE}
#install.packages("tutorial")
library(tutorial)
go_interactive(greedy = FALSE)
```

```{r,tut=T}
# Calculate the average speed of cars. Save as an object called `mean`

# Calculate the standard deviation of the speed of cars. Save as an object called `sd` 


```

## Interactives using Shiny

- Using `runtime: shiny` in the YAML, you can create interactive apps
- I have a problem when making this presentation however... 
- The Shiny runtime does not work with the `tutorial` package!
- My workaround, for this presentation: a separate file

## Interactive Graphs Using Plotly

```{r,echo=F,warning=FALSE,message=FALSE}
library(plotly)
m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)
```

```{r,warning=FALSE,message=FALSE}
p <- plot_ly(cars, x = ~speed, y = ~dist)
p
```

##

```{r}
plot_ly(x = cars$speed, type = "histogram")
```

##

```{r}
plot_ly(x = cars$speed, type = "box")
```


## 2D Histograms 

```{r}
p <- plot_ly(x=cars$speed,y=cars$dist)
pp <- subplot(
  p %>% add_markers(alpha = 0.2),
  p %>% add_histogram2d()
)
pp

```





